### 2018-01-26 analysis_cluster.tex
a first complete pass on clustering method, still needing to be reduced.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_5.pdf}
    \caption{Results of clustering analysis with all 11 attributes (metrics). This results are subsample of 55 possible combination of metrics. Center of cluster is shown as crosses. At this plot ordinal, categorical clusters' numbers are not necessarily represent the poor, fair, good, and excellent clusters. }
    \label{fig:CxCy_11}
\end{figure}
The figure presents a part of results in clustering with 11 features. There are 55 other possibility to present (e.g., C1 vs C7 or C2 vs C9). As one can see in some combination of features the clustering patterns are easily seen. For example in C3 vs C4 we can distinguish 4 region for clusters, although they are partly mixed together. This also correct for C6 vs C7 and C7 vs C8. In the latter one we observe less mixture of data than the former one. Also there are other types of combination where one feature does not add so much information to the clustering process. For example in C9 vs C10, C10 could be considered as irrelevant feature. Irrelevant feature does not add further information for clustering and redundant feature add the equivalent information as the other feature \citep{Dy_2004_MLR}. The last combination of features are a good mixture of data and it is not possible to see the clustering boundaries. C1 vs C2 and C10 vs C11 are examples of such features combinations. It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1, 2, 3, and 4 representing poor, fair, good, and excellent, respectively.\\   
Although we do not expect to see the pattern that resulted from clustering using 11 features in presentation of only two, however, different studies show that the higher dimension reduce the effect of similarity based on distance. \citet{Parsons_2004_ACM} presented an illustrative example to show the effect of dimension in reducing the importance of the distance. Effect of higher dimensions in clustering has been well studied and many different methods are proposed to reduce this effect in the final results. Among them we can name  feature selection before, during, and after clustering, hybrid methods that use combination of methods to select the best subset of feature. \citet{Dy_2004_MLR} addressed two issues involved in developing an automated feature subset selection algorithm for unlabeled data. They illustrated the irrelevant and redundant features and proposed methods for evaluating candidate features using two performance criteria. \\
Subspace analysis is another technique to address the challenges with higher dimensional data in clustering process. Subspace clustering is an extension of traditional clustering which looks for different pattern using subset of features.  \citet{Parsons_2004_ACM} provided a list of algorithm for conducting subspace clustering and also some potential applications for them. The most common factor among these algorithm is the process to find a group of best subspaces through optimization process. An n-dimensional dataset has $2^n$ subspaces where it could be very costly and time consuming to evaluate all of them.\\

@@

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_6.pdf}
    \caption{Box plot of data used in the study after clustering. Metrics are shown for C1 to C11 for 3 components in separate plots for velocity models and clusters. Median of data are shown as notchs. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:C_cvm_comp}
\end{figure}

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_7.pdf}
    \caption{Box plot of data used in the study after clustering and merging the results for different velocity models. Metrics shown for C1 to C11 for 3 components in separate plots for  clusters. Median of data are shown as notchs. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:Figure_cluster_comp}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_8.pdf}
    \caption{Box plot of data used in the study after clustering and merging for different components and velocity models. Metrics shown for C1 to C11  in separate plots for  clusters. Median of data are shown as notchs. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:Figure_cluster_all_comp}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=300px]
        {figures/pdf/Figure_9.pdf}
    \caption{Number of data in each cluster based on components after conducting constraint \kmeans{} clustering through subspace analysis.}
    \label{fig:Figure_cluster_barplot}
\end{figure}

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/Figure_5.pdf}
    \caption{Results of clustering analysis with all 11 attributes (metrics). This results are subsample of 55 possible combination of metrics. Center of cluster is shown as crosses. At this plot ordinal, categorical clusters' numbers are not necessarily represent the poor, fair, good, and excellent clusters. }
    \label{fig:CxCy_11}
\end{figure*}

The figure presents a part of results in clustering with 11 features. There are 55 other possibility to present (e.g., C1 vs C7 or C2 vs C9). As one can see in some combination of features the clustering patterns are easily seen. For example in C3 vs C4 we can distinguish 4 region for clusters, although they are partly mixed together. This also correct for C6 vs C7 and C7 vs C8. In the latter one we observe less mixture of data than the former one. Also there are other types of combination where one feature does not add so much information to the clustering process. For example in C9 vs C10, C10 could be considered as irrelevant feature. Irrelevant feature does not add further information for clustering and redundant feature add the equivalent information as the other feature \citep{Dy_2004_MLR}. The last combination of features are a good mixture of data and it is not possible to see the clustering boundaries. C1 vs C2 and C10 vs C11 are examples of such features combinations. It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1, 2, 3, and 4 representing poor, fair, good, and excellent, respectively.

Although we do not expect to see the pattern that resulted from clustering using 11 features in presentation of only two, however, different studies show that the higher dimension reduce the effect of similarity based on distance. \citet{Parsons_2004_ACM} presented an illustrative example to show the effect of dimension in reducing the importance of the distance. Effect of higher dimensions in clustering has been well studied and many different methods are proposed to reduce this effect in the final results. Among them we can name  feature selection before, during, and after clustering, hybrid methods that use combination of methods to select the best subset of feature. \citet{Dy_2004_MLR} addressed two issues involved in developing an automated feature subset selection algorithm for unlabeled data. They illustrated the irrelevant and redundant features and proposed methods for evaluating candidate features using two performance criteria.

Subspace analysis is another technique to address the challenges with higher dimensional data in clustering process. Subspace clustering is an extension of traditional clustering which looks for different pattern using subset of features.  \citet{Parsons_2004_ACM} provided a list of algorithm for conducting subspace clustering and also some potential applications for them. The most common factor among these algorithm is the process to find a group of best subspaces through optimization process. An n-dimensional dataset has $2^n$ subspaces where it could be very costly and time consuming to evaluate all of them.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/Figure_6.pdf}
    \caption{Box plot of data used in the study after clustering. Metrics are shown for C1 to C11 for 3 components in separate plots for velocity models and clusters. Median of data are shown as notchs. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:C_cvm_comp}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/Figure_7.pdf}
    \caption{Box plot of data used in the study after clustering and merging the results for different velocity models. Metrics shown for C1 to C11 for 3 components in separate plots for  clusters. Median of data are shown as notchs. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:Figure_cluster_comp}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/Figure_8.pdf}
    \caption{Box plot of data used in the study after clustering and merging for different components and velocity models. Metrics shown for C1 to C11  in separate plots for  clusters. Median of data are shown as notchs. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:Figure_cluster_all_comp}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/Figure_9.pdf}
    \caption{Number of data in each cluster based on components after conducting constraint \kmeans{} clustering through subspace analysis.}
    \label{fig:Figure_cluster_barplot}
\end{figure}
